---
title: "City Pairs - International Traffic"
author: "AndrewJ"
date: "4 June 2016"
output: pdf_document
---

# Description

Visualisation experimentation on some random data sets. In this case, it's an [international traffic data](https://data.gov.au/dataset/international-airlines-traffic-by-city-pairs) from data.gov.au, dated 2016-06-02.

# Setup

```{r echo=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(zoo)
library(ggplot2)
library(ggthemes)
library(treemap)
library(formattable)
library(circlize)
```

# Import

```{r}
read_traffic <- function () {
  read_csv("city-pairs-2016-06-02.csv")
}
```

# Transform

```{r}
transform_traffic <- function (df) {
  df %>% 
  mutate(
    Month = as.Date(Month, origin="1899-12-30"),
    AustralianPort = as.factor(AustralianPort),
    ForeignPort = as.factor(ForeignPort),
    Country = as.factor(Country),
    Passengers_In = comma(Passengers_In, digits = 0),
    Passengers_Out = comma(Passengers_Out, digits = 0),
    Passengers_Total = comma(Passengers_Total, digits = 0))
}
```

# Analyse

See the total inbound and outbound traffic by month
```{r}
monthly_traffic <- . %>%
  group_by(Month) %>%
  summarise(
    People_In = sum(Passengers_In), 
    People_Out = sum(Passengers_Out)) %>%
  mutate(
    Roll_Avg_In = rollmean(People_In, k = 12, fill = "extend", align = "center"),
    Roll_Avg_Out = rollmean(People_Out, k = 12, fill = "extend", align = "center"))
```

Get stats on source and destination airport pairs
```{r}
by_airport_pairs <- . %>%
  group_by(AustralianPort, ForeignPort) %>%
  summarise(Passengers_Total = sum(Passengers_Total)) %>%
  ungroup() %>%
  unite(Journey, AustralianPort, ForeignPort, sep = " — ", remove = TRUE)
```

Summarise passengers by country
```{r}
by_country <- . %>%
  group_by(Country) %>%
  summarise(Passengers_Total = sum(Passengers_Total)) %>%
  arrange(desc(Passengers_Total))
```


# Visualise

## Monthly traffic

```{r}
plot_monthly_traffic <- function (df) {
  df %>%
  monthly_traffic() %>%
  gather(key = Direction, value = Passengers,
    People_In, People_Out) %>%
  ggplot(aes(x = Month, y = Passengers, fill = Direction)) +
    geom_bar(stat = 'identity', position = "stack") +
    labs(title = "Monthly Passengers", x = "Month", y = "Passengers") +
    theme_fivethirtyeight()
  }
  
plot_moving_average_traffic <- function (df) {
  df %>%
  monthly_traffic() %>%
  gather(
    key = Direction, value = Passengers, 
    Roll_Avg_In, Roll_Avg_Out) %>%
  ggplot(aes(x=Month, y=Passengers, colour=Direction)) +
    geom_line(stat = 'identity', position = "stack") +
    labs(title = "Moving Average Passengers (k=12)", x = "Month", y = "Passengers") +
    theme_fivethirtyeight()
} 
```

## Plot top countries

```{r}
plot_top_countries <- function (df, n = 10) {
  df %>%
  by_country() %>%
  head(n) %>%
  ggplot(aes(x = reorder(Country, Passengers_Total), y = Passengers_Total)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = Passengers_Total), hjust = 1.0, size = 2.5, colour = "white") +
    coord_flip() +
    labs(title = "Top Countries") +
    theme_fivethirtyeight()
}
```

## Treemap version

```{r}
plot_country_treemap <- function(df, n=25)
  df %>%
  by_country() %>% 
  head(n) %>% 
  treemap(index=c("Country"), vSize=c("Passengers_Total"), title = "Journeys by country")
```

## Plot top n journeys

```{r}
plot_top_journeys <- function (df, n = 10) {
  df %>%
  by_airport_pairs() %>%
  arrange(desc(Passengers_Total)) %>%
  head(n) %>%
  ggplot(aes(x = reorder(Journey, Passengers_Total), y = Passengers_Total)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = Passengers_Total), hjust = 1.1, colour = "white") +
    coord_flip() +
    labs(title = "Top Passenger Journeys", x = "Journey", y = "Passengers") +
    theme_fivethirtyeight()
}
```

# Run

## Import the data.

```{r}
traffic <- read_traffic() %>% transform_traffic()
str(traffic)
levels(traffic$AustralianPort)
levels(traffic$ForeignPort)
```

## Do some plotting

Time series
```{r}
traffic %>% plot_moving_average_traffic()
```

Histograms
```{r}
traffic %>% plot_top_countries(n = 12)
traffic %>% plot_top_journeys(n = 15)
```

Treemap of destinations
```{r}
traffic %>% plot_country_treemap()
```

Chord diagram for top journeys
```{r}
airport_pairs <- traffic %>%
  group_by(AustralianPort, ForeignPort) %>%
  summarise(Passengers = sum(Passengers_Total)) %>%
  ungroup() %>%
  arrange(desc(Passengers)) %>%
  head(20) %>%
  chordDiagram(airport_pairs)
circos.clear()
```


The End